---
title: "CourseProject1 Exploratory Data Analysis"
author: "Pedro alcala"
date: "29/9/2020"
output: html_document
---

## Instructions

This assignment uses data from the UC Irvine Machine Learning Repository, a popular repository for machine learning datasets. In particular, we will be using the “Individual household electric power consumption Data Set” which I have made available on the course web site:

**Dataset:** Electric power consumption [20Mb]
Description: Measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.
The following descriptions of the 9 variables in the dataset are taken from the UCI web site:

* **Date:** Date in format dd/mm/yyyy
* **Time:** time in format hh:mm:ss
* **Global_active_power:** household global minute-averaged active power (in kilowatt)
* **Global_reactive_power:** household global minute-averaged reactive power (in kilowatt)
* **Voltage:** minute-averaged voltage (in volt)
* **Global_intensity:** household global minute-averaged current intensity (in ampere)
* **Sub_metering_1:** energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
* **Sub_metering_2:** energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
* **Sub_metering_3:** energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.

## Plot 1

```{r}
library("data.table")

setwd("C:\\Users\\lenovo\\Documents\\Data Science Coursera\\Curso 4")

#Reads in data from file then subsets data for specified dates
powerDT <- data.table::fread(input = "household_power_consumption.txt"
                             , na.strings="?"
)

# Prevents histogram from printing in scientific notation
powerDT[, Global_active_power := lapply(.SD, as.numeric), .SDcols = c("Global_active_power")]

# Change Date Column to Date Type
powerDT[, Date := lapply(.SD, as.Date, "%d/%m/%Y"), .SDcols = c("Date")]

# Filter Dates for 2007-02-01 and 2007-02-02
powerDT <- powerDT[(Date >= "2007-02-01") & (Date <= "2007-02-02")]

png("plot1.png", width=480, height=480)

## Plot 1
hist(powerDT[, Global_active_power], main="Global Active Power", 
     xlab="Global Active Power (kilowatts)", ylab="Frequency", col="Red")

dev.off()


```


## Plot 2

```{r}
library("data.table")

setwd("C:\\Users\\lenovo\\Documents\\Data Science Coursera\\Curso 4")

#Reads in data from file then subsets data for specified dates
powerDT <- data.table::fread(input = "household_power_consumption.txt"
                             , na.strings="?"
)

# Prevents Scientific Notation
powerDT[, Global_active_power := lapply(.SD, as.numeric), .SDcols = c("Global_active_power")]

# Making a POSIXct date capable of being filtered and graphed by time of day
powerDT[, dateTime := as.POSIXct(paste(Date, Time), format = "%d/%m/%Y %H:%M:%S")]

# Filter Dates for 2007-02-01 and 2007-02-02
powerDT <- powerDT[(dateTime >= "2007-02-01") & (dateTime < "2007-02-03")]

png("plot2.png", width=480, height=480)

## Plot 2
plot(x = powerDT[, dateTime]
     , y = powerDT[, Global_active_power]
     , type="l", xlab="", ylab="Global Active Power (kilowatts)")

dev.off()


```



## Plot 3

```{r}
library("data.table")

setwd("C:\\Users\\lenovo\\Documents\\Data Science Coursera\\Curso 4")

#Reads in data from file then subsets data for specified dates
powerDT <- data.table::fread(input = "household_power_consumption.txt"
                             , na.strings="?"
)

# Prevents Scientific Notation
powerDT[, Global_active_power := lapply(.SD, as.numeric), .SDcols = c("Global_active_power")]

# Making a POSIXct date capable of being filtered and graphed by time of day
powerDT[, dateTime := as.POSIXct(paste(Date, Time), format = "%d/%m/%Y %H:%M:%S")]

# Filter Dates for 2007-02-01 and 2007-02-02
powerDT <- powerDT[(dateTime >= "2007-02-01") & (dateTime < "2007-02-03")]

png("plot3.png", width=480, height=480)

# Plot 3
plot(powerDT[, dateTime], powerDT[, Sub_metering_1], type="l", xlab="", ylab="Energy sub metering")
lines(powerDT[, dateTime], powerDT[, Sub_metering_2],col="red")
lines(powerDT[, dateTime], powerDT[, Sub_metering_3],col="blue")
legend("topright"
       , col=c("black","red","blue")
       , c("Sub_metering_1  ","Sub_metering_2  ", "Sub_metering_3  ")
       ,lty=c(1,1), lwd=c(1,1))

dev.off()


```


## Plot 4

```{r}

library("data.table")

setwd("C:\\Users\\lenovo\\Documents\\Data Science Coursera\\Curso 4")

#Reads in data from file then subsets data for specified dates
powerDT <- data.table::fread(input = "household_power_consumption.txt"
                             , na.strings="?"
)

# Prevents Scientific Notation
powerDT[, Global_active_power := lapply(.SD, as.numeric), .SDcols = c("Global_active_power")]

# Making a POSIXct date capable of being filtered and graphed by time of day
powerDT[, dateTime := as.POSIXct(paste(Date, Time), format = "%d/%m/%Y %H:%M:%S")]

# Filter Dates for 2007-02-01 and 2007-02-02
powerDT <- powerDT[(dateTime >= "2007-02-01") & (dateTime < "2007-02-03")]

png("plot4.png", width=480, height=480)

par(mfrow=c(2,2))

# Plot 1
plot(powerDT[, dateTime], powerDT[, Global_active_power], type="l", xlab="", ylab="Global Active Power")

# Plot 2
plot(powerDT[, dateTime],powerDT[, Voltage], type="l", xlab="datetime", ylab="Voltage")

# Plot 3
plot(powerDT[, dateTime], powerDT[, Sub_metering_1], type="l", xlab="", ylab="Energy sub metering")
lines(powerDT[, dateTime], powerDT[, Sub_metering_2], col="red")
lines(powerDT[, dateTime], powerDT[, Sub_metering_3],col="blue")
legend("topright", col=c("black","red","blue")
       , c("Sub_metering_1  ","Sub_metering_2  ", "Sub_metering_3  ")
       , lty=c(1,1)
       , bty="n"
       , cex=.5) 

# Plot 4
plot(powerDT[, dateTime], powerDT[,Global_reactive_power], type="l", xlab="datetime", ylab="Global_reactive_power")

dev.off()
```

